Computational Prediction of Boundaries of 3D Genomic Domains in Class Imbalance

### 2019-04-24 Manuscript/manuscript.Rmd
manuscript edits

<!--
The high correlations among many of the genomic annotations gave precedent for the use of an elastic-net regularization algorithm in order to suitably reduce the feature space while retaining possibly correlated annotations that could be predictive of TAD boundary formation. 
-->

@@

@@

Distance profiles on a $log_{2}$ scale were found to exhibit significantly different densities between classes of genomic bins for TFBS with highly ranked predictive importances. Using permutation tests we found that the transcription factors SMC3a, ZNF143, CTCF, and RAD21 tended to have binding peaks that that were located closer to the centers of 10 kb sized genomic bins that overlapped with flanked TAD boundaries than to bins that did not overlap with flanked TAD boundaries across both cell lines (Figure 7 A-H). Similar distance profiles for the same TFBS were obtained when assessed at lower resolutions, across both cell lines (Supplementary Figures 12 & 13 A-L).

>>>>

The high correlations among many of the genomic annotations gave precedent for the use of an elastic-net regularization algorithm in order to suitably reduce the feature space while retaining possibly correlated annotations that could be predictive of TAD boundary formation. For each model created from the ensemble framework, the alpha and lambda were tuned over a grid of values using 10-fold cross validation (Supplementary Table 7).

@@

## MODELS IMPLEMENTED ON TRANSCRIPTION FACTOR BINDING SITES OUTPERFORM OTHER ANNOTATION-SPECIFIC MODELS

@@

## DISTANCE PROFILES HIGHLIGHT THE RELATIONSHIP BETWEEN KEY TFBS AND TAD BOUNDARIES

When evaluating the 10 kb data resolution, on GM12878, the distance profiles (on a $log_{2}$ scale) were found to be smaller for top predictive TFBS annotations in relation to genomic bins that overlapped with flanked TAD boundaries, when compared to bins that did not overlap with a flanked TAD boundary (Figure 7). At the same resolution and cell line, direct comparisons of quantiles obtained from the distribution of distances for the transcription factors SMC3a, ZNF143, CTCF, and RAD21, led to the the conclusion that these TFBS were more often located near TAD boundaries on the linear genome, as can be seen from the shifted quantile differences (Figure 7). Similar distance profiles for the same TFBS were obtained when assessed at lower resolutions, across both cell lines (Supplementary Figures 12-17 A-D).
